Human capital mediates natural selection in contemporary humans

### 2022-01-02 why-natural-selection.Rmd
Run a single regression on all 29 scores simultaneousl within-siblings

@@

lm_pool_w <- tidy(lm(estimate.y ~ estimate.x, res_all))
coef_pool_w <- lm_pool_w[[2, "estimate"]]
se_pool_w <- lm_pool_w[[2, "std.error"]]
```

These results support our theory that natural selection on polygenic scores is
driven by their correlation with human capital. This need not imply
that a given polygenic score is causal of either fertility or human capital. Instead
scores may simply correlate with causally relevant environments, for example via
"genetic nurture" effects [@kong2018nature] or population stratification. To 
investigate the presence of these indirect genetic effects, we estimate equation
\@ref(eq:regression) among full siblings, controlling for a full set of sibling
group dummies (Appendix Figure \@ref(fig:plot-res-fe-fertility)). With a reduced
sample size, all within-sibling effects are insignificant after Bonferroni
correction. However, effect sizes are positively correlated with effect sizes
from the pooled model, and about 1/3 smaller (regressing within-sibling on
pooled effect sizes, $b$ = `r coef_pool_w`). This attenuation is broadly in line
with the ~50% decrease in heritability in within-sibling GWASs on age at first
birth and educational attainment [@Howe2021.03.05.433935]. We see these results
as providing tentative evidence that polygenic scores cause fertility, with effects
being at least partly driven by gene-environment correlations. We also reran
within-siblings regressions adding a control for education. Most effect sizes
barely change, suggesting that our measure of education does not in general
mediate differences in fertility among siblings. However, for educational
attainment polygenic scores, effect sizes are reduced by about 20%.

@@

```{r plot-res-fe-fertility, fig.cap = "Selection effects controlling for sibling-group fixed effects, with and without a control for education (left education before 16, 16-18, or after 18)."}

drake::loadd(res_fe_fertility)

standard_ggplot(res_fe_fertility, fill_col = Regression, n_regs = 33, fill_direction = -1)

@@

>>>>

educ_reduction <- with(res_fe_fertility, 
       estimate[Regression == "Controlled" & term == "EA3_excl_23andMe_UK"] / 
       estimate[Regression == "Raw" & term == "EA3_excl_23andMe_UK"]
     )
educ_reduction <- 1 - educ_reduction

@@

lm_pool_w <- tidy(lm(estimate.y ~ estimate.x, res_all))
coef_pool_w <- lm_pool_w[[2, "estimate"]]
se_pool_w <- lm_pool_w[[2, "std.error"]]

These results support our theory that natural selection on polygenic scores is
driven by their correlation with human capital. We also test for causality by
running within-siblings regressions (Appendix Figure
\@ref(fig:plot-res-fe-fertility)). With a reduced sample size, all
within-sibling effects are insignificant after Bonferroni correction. However,
effect sizes are positively correlated with effect sizes from the pooled model,
and about 50% smaller (regressing within-sibling on pooled effect sizes, $b$ =
`r format(coef_pool_w, digits = 3)`). This attenuation is in line with the ~50%
decrease in heritability in within-sibling GWASs on age at first birth and
educational attainment [@Howe2021.03.05.433935]. We see these results as
providing tentative evidence that polygenic scores cause fertility, with effects
being partly driven by gene-environment correlations. We also reran
within-siblings regressions adding a control for education. Most effect sizes
barely change, suggesting that our measure of education does not in general
mediate differences in fertility among siblings. However, for educational
attainment polygenic scores, effect sizes are reduced by `r percent(educ_reduction)`.

@@

```{r plot-res-fe-fertility, fig.cap = "Selection effects controlling for sibling-group fixed effects, with and without a control for education (left education before 16, 16-18, or after 18). Each set of results is from a single regression of 'n children' on 29 polygenic scores. Standard errors clustered by sibling group."}

drake::loadd(res_fe_fertility)
n_regs <- nrow(res_fe_fertility)/2 # Raw and Controlled treated separately

standard_ggplot(res_fe_fertility, fill_col = Regression, n_regs = n_regs,
                  fill_direction = -1)

n_groups <- attr(res_fe_fertility, "n_groups")
n <- attr(res_fe_fertility, "n")

@@

We run a single regression on 29 polygenic scores within `r n_groups` sibling
groups (N = `r n`). Thus, we control both for environmental confounds (since
scores are randomly allocated within sib-groups by meiosis), and for genetic
confounds captured by our polygenic scores. We remove four scores which
correlate highly with other scores (educational attainment 2, hip circumference,
waist circumference and waist-hip ratio). Figure \@ref(fig:plot-res-fe-fertility)
shows the results.
